---
title: "Forecasting Dengue Spread"
author: "Srihari Seshadri, Siddhartha Khetawat"
date: "June 3, 2019"
output: html_document
---

```{r message=FALSE, warning=FALSE}
# Load libraries and install them, if necessary
tmp.library.list <- c("haven", "zoo", "fUnitRoots", "tseries", "urca", "lmtest", "forecast", "data.table", "readxl","reshape", "quantmod", "ggplot2", "reshape2", "plyr","scales", "hts", "fpp2", "lubridate","stargazer","GGally", "dplyr", "lattice", "skimr")
for (i in 1:length(tmp.library.list)) {
  if (!tmp.library.list[i] %in% rownames(installed.packages())) {
    install.packages(tmp.library.list[i])
  }
  library(tmp.library.list[i], character.only = TRUE)
}
rm(tmp.library.list)
```


# Loading the Data

```{r}
trainraw <- data.table(read.csv("dengue_features.csv"), sep=',', stringsAsFactors = F)
trainlraw <- data.table(read.csv("dengue_labels.csv"), sep=',', stringsAsFactors = F)

head(trainraw)
head(trainlraw)

dim(trainraw)
dim(trainlraw)
```


Merge the features data with the labels data

```{r}
sjdata <- merge(trainraw, trainlraw, by=c("city", "year", "weekofyear"), all.x = T)
nrow(sjdata) == nrow(trainraw) # if true, we did not accidentally merge many-to-one 

dim(sjdata)
```

Omit reanalysis_sat_precip_amt_mm variable because of large amount of missing data

```{r message=FALSE, warning=FALSE}
sjdata[, reanalysis_sat_precip_amt_mm := NULL]
```

Sort the dataset so it's in a clear order

```{r}
setkeyv(sjdata, c("city", "year", "weekofyear"))

skim(sjdata)
```

```{r}
head(sjdata)
```


# Splitting the data into training and testing datasets

Transform the Date column into the appropriate type
```{r}
sjdata[,week_start_date := as.Date(week_start_date,format = "%m/%d/%Y")]
```


Look at date ranges
```{r}
min(sjdata$week_start_date)
max(sjdata$week_start_date)
```

```{r}
# Preserve 5 years of data for model validation
sjdata.train <- sjdata[week_start_date < "2003-01-01"]
sjdata.test <- sjdata[week_start_date >= "2003-01-01"]
```

# Descriptive analytics and data visualisation


```{r}
head(sjdata.train)

# Summary
summary(sjdata.train)

# Correlation plots
ggcorr(sjdata.train[, 5:26], label = TRUE, hjust = 0.85, size = 2,color = "grey50",
       label_size = 2) +
  ggplot2::labs(title = "Correlation Plot")

ggsave("./Charts/EDA_correlation.png", width = 8, height = 6)

```

```{r message=FALSE, warning=FALSE}
ggplot(data=sjdata.train, aes(sjdata.train$total_cases)) + 
  geom_histogram() + 
  ggtitle("Distrbution of total cases of Dengue") 

ggsave("./Charts/total_cases_distribution.png", width = 8, height = 6)

```

```{r}
(missing_idxs <- c(which(is.na(sjdata.train$total_cases), arr.ind=TRUE)))
sjdata.train[missing_idxs,] %>% select(year, weekofyear, total_cases)
```


```{r}
# Plot total cases with timeline -> seasonal effect
ggplot(sjdata.train , aes(week_start_date, total_cases)) + 
  geom_line() + 
  scale_y_continuous() +
  scale_x_date(breaks = date_breaks("year"), labels = date_format("%m-%Y")) +
  ylab("Total cases in San Juan") +
  xlab("")+
  theme(axis.text.x=element_text(angle=-70, hjust=0.001)) +
  labs(title = "Time Series Plot of Total Cases") +
  theme_bw()

ggsave("./Charts/time_series_totalcases.png", width = 8, height = 6)

```

```{r}
# Scatter Plot Total cases vs. Max Temperature
plot(total_cases ~ station_max_temp_c, data=sjdata.train
     ,main = "Scatter Plot of Total Cases and Maximum Temperature in San Juan")
```

```{r}
# Scatter Plot Total cases vs. Humidity
plot(total_cases ~ reanalysis_specific_humidity_g_per_kg, data=sjdata.train,
     main = "Scatter Plot of Total Cases and Humidity in San Juan")
```

```{r}
# Time Series Plot of log(Total cases) & Max Temperature
ggplot(sjdata.train, aes(week_start_date)) +
  geom_line(aes(y = station_max_temp_c, colour = "station_max_temp_c")) +
  geom_line(aes(y = (y = log(total_cases+1)), 
                colour = "log total_cases")) +
  labs(title = "Log Total Cases and Maximum Temperature ") +
  theme_bw() +theme(legend.position="bottom")

ggsave("./Charts/time_series_logtotalcases_maxtemp.png", width = 8, height = 6)

```

```{r}
# Time Series Plot of log(Total cases) & Humidity
ggplot(sjdata.train, aes(week_start_date)) +
  geom_line(aes(y = log(total_cases+1), 
                colour = "log total_cases")) +
  geom_line(aes(y = reanalysis_specific_humidity_g_per_kg, 
                colour = "reanalysis_specific _humidity_g_per_kg")) +
  labs(title = "Log Total Cases and Humidity") +
  theme_bw() + theme(legend.position="bottom")

ggsave("./Charts/time_series_logtotalcases_humidity.png", width = 8, height = 6)
```


# Data Imputation

```{r}
# Build a function to impute data with the most recent that is non missing (Using LOCF)
na.locf.data.frame <- 
  function(object, ...) replace(object, TRUE, lapply(object, na.locf, ...))

# Fill in NAs
sjdata.train.imputed <- na.locf.data.frame(sjdata.train)
summary(sjdata.train.imputed)

sjdata.train <- sjdata.train.imputed
```

# Diagnostic study

```{r}
# Convert data to TS with weekly frequency
sjtrain.ts <- ts(sjdata.train$total_cases, frequency = 52, start = c(1990,18))
head(sjtrain.ts)
```

Stationarity test

```{r}
# ADF Test
adf.test(sjdata.train$total_cases)

# KPSS Test
kpss.test(sjdata.train$total_cases)
```

The time series is stationary.

```{r}
# ACF Plot
ggAcf(sjtrain.ts) +
  labs(title = "Autocorrelation Function Plot") +
  theme_bw()

ggsave("./Charts/ACF.png", width = 8, height = 6)


# PACF Plot
ggPacf(sjtrain.ts)  +
  labs(title = "Partial Autocorrelation Function Plot (San Juan)") +
  theme_bw()

ggsave("./Charts/PACF.png", width = 8, height = 6)

```

We can observe significant autocorrelation and seasonality.

Decompose the time series 

```{r}
decomp = stl(sjtrain.ts, s.window="periodic")
plot(decomp)
```

# Modeling

The following models were used for forecasting dengue spread.

## ARIMA

Use ARIMA without seasonality on the training data and analyse the residuals.

```{r}
sj.fit1 <- auto.arima(sjtrain.ts, seasonal = F)
sj.fit1

checkresiduals(sj.fit1)
```


## SARIMA

Apply Seasonal ARIMA to the data and analyse the residuals.

```{r}
sj.fit2 <- auto.arima(sjtrain.ts, seasonal = T)
sj.fit2

checkresiduals(sj.fit2)
```


## SARIMAX

```{r, warning=FALSE}
# Prepare xreg
sjtrain.ts1 <- ts(sjdata.train.imputed,
                  freq=365.25/7,
                  start=decimal_date(ymd("1990-04-30")))

varlist <- c("precipitation_amt_mm"
             ,"reanalysis_dew_point_temp_k"
             ,"reanalysis_relative_humidity_percent"
             ,"station_avg_temp_c"
             ,"station_diur_temp_rng_c"
             ,"station_max_temp_c"
             ,"station_min_temp_c")

create.tslag <- function(x, dataset) {
  cbind(
    Lag0 = dataset[,x],
    Lag1 = stats::lag(dataset[,x],-1),
    Lag2 = stats::lag(dataset[,x],-2),
    Lag3 = stats::lag(dataset[,x],-3),
    Lag4 = stats::lag(dataset[,x],-4)) %>%
    head(NROW(dataset)) 
}

```


Plot distributions of the feature vectors


```{r message=FALSE, warning=FALSE}
ggplot() + 
  geom_line(mapping=aes(x=week_start_date, y=precipitation_amt_mm), data=sjdata.train.imputed) +
  theme_bw() + 
  ggtitle("Precipitation across time")

ggplot() + 
  geom_line(mapping=aes(x=week_start_date, y=station_avg_temp_c), data=sjdata.train.imputed) +
  theme_bw() + 
  ggtitle("Average temperature across time")

ggplot() + 
  geom_line(mapping=aes(x=week_start_date, y=station_min_temp_c), data=sjdata.train.imputed) +
  theme_bw() + 
  ggtitle("Miimum temperature across time")

ggplot() + 
  geom_line(mapping=aes(x=week_start_date, y=reanalysis_dew_point_temp_k), data=sjdata.train.imputed) +
  theme_bw() + 
  ggtitle("Dew point across time")


ggplot() + 
  geom_line(mapping=aes(x=week_start_date, y=station_diur_temp_rng_c), data=sjdata.train.imputed) +
  theme_bw() + 
  ggtitle("Diur temperature across time")

ggplot() + 
  geom_line(mapping=aes(x=week_start_date, y=reanalysis_relative_humidity_percent), data=sjdata.train.imputed) +
  theme_bw() + 
  ggtitle("Relative humidity across time")

ggplot() + 
  geom_line(mapping=aes(x=week_start_date, y=station_max_temp_c), data=sjdata.train.imputed) +
  theme_bw() + 
  ggtitle("Station max temperature across time")

```


```{r}
precipitation <- create.tslag(x = "precipitation_amt_mm"
                              ,sjtrain.ts1)

dew.temp <- create.tslag(x = "reanalysis_dew_point_temp_k"
                         ,sjtrain.ts1)

relative.humidity <- create.tslag(x = "reanalysis_relative_humidity_percent"
                                  ,sjtrain.ts1)

avg.temp <- create.tslag(x = "station_avg_temp_c"
                         ,sjtrain.ts1)

diur.temp <- create.tslag(x = "station_diur_temp_rng_c"
                          ,sjtrain.ts1)

max.temp <- create.tslag(x = "station_max_temp_c"
                         ,sjtrain.ts1)

min.temp <- create.tslag(x = "station_min_temp_c"
                         ,sjtrain.ts1)
```

Apply Seasonal ARIMAX to the data and analyse the residuals.

```{r}
sj.fit3 <- auto.arima(sjtrain.ts, xreg=cbind(precipitation[,1],relative.humidity[,1]
                                          ,dew.temp[,1],avg.temp[,1]
                                          ,diur.temp[,1],max.temp[,1]
                                          ,min.temp[,1]), seasonal=TRUE)
sj.fit3
checkresiduals(sj.fit3)
```

## Neural Network

```{r}
# Fit Neural Network
sj.fit4 <- nnetar(sjtrain.ts,xreg=sjtrain.ts1[,5:23])

sj.fit4
checkresiduals(sj.fit4)
```

```{r}
Box.test(sj.fit4$residuals, type = "Ljung-Box")
```


# Model Selection

```{r, warning=FALSE}
# Fill in NAs
sjdata.test.imputed <- na.locf.data.frame(sjdata.test)
summary(sjdata.test.imputed)

# Convert data to TS
sjtest.ts1 <- ts(sjdata.test.imputed, 
                 freq=52, 
                 start=decimal_date(ymd("2003-01-01")))
```

```{r}
# Take lags of predictors
precipitation1 <- create.tslag(x = "precipitation_amt_mm"
                               ,sjtest.ts1)

dew.temp1 <- create.tslag(x = "reanalysis_dew_point_temp_k"
                          ,sjtest.ts1)

relative.humidity1 <- create.tslag(x = "reanalysis_relative_humidity_percent"
                                   ,sjtest.ts1)

avg.temp1 <- create.tslag(x = "station_avg_temp_c"
                          ,sjtest.ts1)

diur.temp1 <- create.tslag(x = "station_diur_temp_rng_c"
                           ,sjtest.ts1)

max.temp1 <- create.tslag(x = "station_max_temp_c"
                          ,sjtest.ts1)

min.temp1 <- create.tslag(x = "station_min_temp_c"
                          ,sjtest.ts1)

```

## Evaluate the model accuracies

```{r}
# Model 1 (ARIMA)
fc0 <- forecast(sj.fit1,h=277)
plot(fc0)
accuracy(fc0$mean, as.integer(sjtest.ts1[,"total_cases"]))

ggplot() + 
  geom_line(mapping=aes(x=week_start_date, y=total_cases, color="ori"), data=sjdata.test) + 
  geom_line(mapping=aes(x=sjdata.test$week_start_date, y=fc0$mean, color="pred")) + 
  theme_light() + theme(legend.position = "right") + 
  ggtitle("ARIMA vs. Ground Truth")

ggsave("./Charts/ARIMA vs. Ground Truth.png", width = 8, height = 6)

```

```{r}
# Model 2 (SARIMA)
fc1 <- forecast(sj.fit2, h=277)
plot(fc1)
accuracy(fc1$mean, as.integer(sjtest.ts1[,"total_cases"]))

ggplot() + 
  geom_line(mapping=aes(x=week_start_date, y=total_cases, color="ori"), data=sjdata.test) + 
  geom_line(mapping=aes(x=sjdata.test$week_start_date, y=fc1$mean, color="pred")) + 
  theme_light() + theme(legend.position = "right") + 
  ggtitle("SARIMA vs. Ground Truth")

ggsave("./Charts/SARIMA vs. Ground Truth.png", width = 8, height = 6)

```


```{r, warning=FALSE}
# Model 3 (SARIMAX)
fc3 <- forecast(sj.fit3, xreg=cbind(precipitation1[,1],relative.humidity1[,1]
                                 ,dew.temp1[,1],avg.temp1[,1]
                                 ,diur.temp1[,1],max.temp1[,1]
                                 ,min.temp1[,1]), h =277)
plot(fc3)
accuracy(fc3$mean,as.integer(sjtest.ts1[,"total_cases"]))

ggplot() + 
  geom_line(mapping=aes(x=week_start_date, y=total_cases, color="ori"), data=sjdata.test) + 
  geom_line(mapping=aes(x=sjdata.test$week_start_date, y=fc3$mean, color="pred")) + 
  theme_light() + theme(legend.position = "right") + 
  ggtitle("SARIMAX vs. Ground Truth")

ggsave("./Charts/SARIMAX vs. Ground Truth.png", width = 8, height = 6)

```


```{r}
# Try NN
fc4 <- forecast(sj.fit4, xreg=sjtest.ts1[,5:23], h=277)
autoplot(fc4)
accuracy(fc4$mean,as.integer(sjtest.ts1[,"total_cases"]))

ggplot() + 
  geom_line(mapping=aes(x=week_start_date, y=total_cases, color="ori"), data=sjdata.test) + 
  geom_line(mapping=aes(x=sjdata.test$week_start_date, y=fc4$mean, color="pred")) + 
  theme_light() + theme(legend.position = "right") + 
  ggtitle("Neural networks prediction vs. Ground Truth")

ggsave("./Charts/Neural networks prediction vs. Ground Truth.png", width = 8, height = 6)

```

```{r}

```

